---
title: "Homework Assignment #4"
subtitle: "Prioritizing potential aquaculture"
author: "Haylee Oyler"
date: last-modified
execute: 
  eval: true
  warning: false
  message: false
format:
  html:
    toc: true
editor_options: 
  chunk_output_type: console
---

### Load packages

```{r}
#| code-fold: true
#| code-summary: "Reveal code"

library(sf)
library(here)
library(terra)
library(tmap)
library(tidyverse)
library(viridisLite)
library(RColorBrewer)
library(kableExtra)
library(testthat)
library(patchwork)
```

## Read in data

```{r}
west_eez <- read_sf(here("data", "wc_regions_clean.shp"))
depth <- terra::rast(here("data", "depth.tif"))
```

```{r}
# Compile all our SST rasters into one object
sst_files <- list.files(here("data"), 
                        pattern = "average", 
                        full.names = TRUE)

```

```{r}
# Generate an empty vector 
# sst <- c()
# 
# # Fill the vector with out sst files
# for (i in seq_along(sst_files)) {
#   sst_file <- terra::rast(sst_files[i])
#   sst[i] <- sst_file
# }
```

```{r}
sst <- c(terra::rast(here("data", "average_annual_sst_2008.tif")),
         terra::rast(here("data", "average_annual_sst_2009.tif")),
         terra::rast(here("data", "average_annual_sst_2010.tif")),
         terra::rast(here("data", "average_annual_sst_2011.tif")),
         terra::rast(here("data", "average_annual_sst_2012.tif"))
         )
```

```{r}
st_crs(west_eez) == st_crs(depth)
st_crs(sst) == st_crs(depth)

st_crs(west_eez)
st_crs(depth)
st_crs(sst)

```

### Check that CRS match

```{r}
#| warning: true
#| code-fold: true
#| code-summary: "Reveal code"

data_list <- list(depth, sst, west_eez)

for (i in 1:length(data_list)) {
  if (st_crs(data_list[[i]]) != st_crs(depth)) {
 warning("The CRS of do not match, transforming now...")
  data_list[[i]] <- terra::project(data_list[[i]], depth) 
 } else {
    print(paste("The CRS match"))
  }
}
```


## Process data
find the mean SST from 2008-2012 (e.g. create single raster of average SST)
```{r}
mean_sst <- app(sst, fun = mean)
```

convert average SST from Kelvin to Celsius
hint: subtract by 273.15
```{r}
mean_sst <- mean_sst - 273.15
```

crop depth raster to match the extent of the SST raster
note: the resolutions of the SST and depth data do not match
```{r}
depth_crop <- crop(depth, sst)
```

resample the depth data to match the resolution of the SST data using the nearest neighbor approach
```{r}
depth_resample <- resample(depth_crop, y = sst, method = "near")
```

```{r}
# depth_resample <- st_transform(depth_resample, crs = st_crs(sst)) 
depth_resample <- project(depth_resample, crs(sst))
```

check that the depth and SST match in resolution, extent, and coordinate reference system
hint: can the rasters be stacked?
```{r}
depth_sst <- c(depth_resample, sst)
```


## Find suitable locations
reclassify SST and depth data into locations that are suitable for oysters
hint: set suitable values to 1 and unsuitable values to 0
```{r}

```

find locations that satisfy both SST and depth conditions
```{r}

```

## Determine the most suitable EEZ
We want to determine the total suitable area within each EEZ in order to rank zones by priority. To do so, we need to find the total area of suitable locations within each EEZ.

select suitable cells within West Coast EEZs
```{r}

```

find area of grid cells
```{r}

```

find the total suitable area within each EEZ
hint: it might be helpful to rasterize the EEZ data
```{r}

```

