---
title: "Homework Assignment #4"
subtitle: "Prioritizing potential aquaculture"
author: "Haylee Oyler"
date: last-modified
execute: 
  eval: true
  warning: false
  message: false
format:
  html:
    toc: true
editor_options: 
  chunk_output_type: console
---

### Load packages

```{r}
#| code-fold: true
#| code-summary: "Reveal code"

library(sf)
library(here)
library(terra)
library(tmap)
library(tidyverse)
library(viridisLite)
library(RColorBrewer)
library(kableExtra)
library(testthat)
library(patchwork)
```

## Read in data

```{r}
west_eez <- read_sf(here("data", "wc_regions_clean.shp"))
depth <- terra::rast(here("data", "depth.tif"))
```

```{r}
sst2008 <- terra::rast(here("data", "average_annual_sst_2008.tif"))
sst2009 <- terra::rast(here("data", "average_annual_sst_2009.tif"))
sst2010 <- terra::rast(here("data", "average_annual_sst_2010.tif"))
sst2011 <- terra::rast(here("data", "average_annual_sst_2011.tif"))
sst2012 <- terra::rast(here("data", "average_annual_sst_2012.tif"))

sst <- c(sst2008, sst2009, sst2010, sst2011, sst2012)
```


```{r}
# Compile all our SST rasters into one object
sst_files <- list.files(here("data"), 
                        pattern = "average", 
                        full.names = TRUE)

```

```{r}
# Generate an empty vector 
# sst <- vector(length = length(sst_files))
# # Vector has no length so it won't fill, need to specify how long you want it to be
# 
# # Fill the vector with sst files
# for (i in seq_along(sst_files)) {
#   sst_file <- terra::rast(sst_files[i])
#   sst[i] <- sst_file
# }
```

```{r}
# sst <- c(terra::rast(here("data", "average_annual_sst_2008.tif")),
#          terra::rast(here("data", "average_annual_sst_2009.tif")),
#          terra::rast(here("data", "average_annual_sst_2010.tif")),
#          terra::rast(here("data", "average_annual_sst_2011.tif")),
#          terra::rast(here("data", "average_annual_sst_2012.tif"))
#          )
```


```{r}
#| code-fold: true
#| code-summary: "Reveal code"

if(st_crs(west_eez) == st_crs(depth)) {
  print("Coordinate reference systems match")
} else{
  warning("Updating coordinate reference systems to match")
  # transform data to match
  west_eez <- st_transform(west_eez, st_crs(depth))
}

if(st_crs(sst) == st_crs(depth)) {
  print("Coordinate reference systems match")
} else{
  warning("Updating coordinate reference systems to match")
  # transform data to match
  sst <- project(sst, crs(depth))
}
```


```{r}
st_crs(west_eez) == st_crs(depth)
st_crs(sst) == st_crs(depth)

# st_crs(west_eez)
# st_crs(depth)
# st_crs(sst)
```

Pacific rock crabs
depth range 0 - 91 m

Pacific sea star
depth range 0 - 30 m 
24.7 - 29.3, mean 28.5

### Check that CRS match

```{r}
#| warning: true
#| code-fold: true
#| code-summary: "Reveal code"

# data_list <- list(depth, sst, west_eez)
# 
# for (i in 1:length(data_list)) {
#   if (st_crs(data_list[[i]]) != st_crs(depth)) {
#  warning("The CRS of do not match, transforming now...")
#   data_list[[i]] <- terra::project(data_list[[i]], depth) 
#  } else {
#     print(paste("The CRS match"))
#   }
# }

```


## Process data
find the mean SST from 2008-2012 (e.g. create single raster of average SST)
convert average SST from Kelvin to Celsius

```{r}
# depth <- project(depth, crs("EPSG:4326"))
# sst <- project(sst, crs("EPSG:4326"))

mean_sst <- app(sst, fun = mean)

mean_sst <- mean_sst - 273.15
```

```{r}
# Crop the depth raster to match extent of sst
depth_crop <- crop(depth, mean_sst)

# Resample depth to match resolution of sst
depth_resample <- resample(depth_crop, y = mean_sst, method = "near")

```

check that the depth and SST match in resolution, extent, and coordinate reference system
hint: can the rasters be stacked?
```{r}
test_stack <- c(depth_resample, mean_sst)
```


## Find suitable locations
reclassify SST and depth data into locations that are suitable for oysters
hint: set suitable values to 1 and unsuitable values to 0
sea surface temperature: 11-30°C
depth: 0-70 meters below sea level

```{r}
# create sst reclassification matrix
rcl_sst <- matrix(c(-Inf, 11, NA, 
                    11, 30, 1, 
                    30, Inf, NA),
                    ncol = 3, byrow = TRUE)

# use reclassification matrix to reclassify sst raster
reclass_sst <- classify(mean_sst, rcl = rcl_sst)

plot(reclass_sst)
```


```{r}
# create sst reclassification matrix
rcl_depth <- matrix(c(-Inf, -70, NA, 
                    -70, 0, 1, 
                    0, Inf, NA),
                    ncol = 3, byrow = TRUE)

# use reclassification matrix to reclassify sst raster
reclass_depth <- classify(depth_resample, rcl = rcl_depth)

plot(reclass_depth)
# change reclassified values into factors
# values(reclassified_depth) <- as.factor(values(reclassified_depth))
```

find locations that satisfy both SST and depth conditions
```{r}
sst_depth <- lapp(c(reclass_sst, reclass_depth), fun = "*")
# sst_depth <- (reclass_depth == 1) & (reclass_sst == 1)

# sst_depth <- reclass_sst * reclass_depth
plot(sst_depth)

```

```{r}
suitable_area <- cellSize(x = sst_depth, 
                          mask = TRUE,
                          unit = 'km') # Selecting km from data

# Initial plot
plot(suitable_area)
# Find the total suitable area within each EEZ

# Rasterize EEZ data
eez_raster <- rasterize(west_eez,
                        suitable_area, # to this raster
                        field = 'rgn') # Transfer values to each eez region

# Initial plot
plot(eez_raster)

tm_shape(eez_raster) +
  tm_raster() +
  tm_shape(sst_depth) +
  tm_raster()

# Use zonal algebra to aggregate a grouping variable
eez_suitable <- zonal(x = suitable_area, 
                      z = eez_raster, # Raster representing zones
                      fun = 'sum', # To add up total area
                      na.rm = TRUE)
eez_suitable
```


## Determine the most suitable EEZ
We want to determine the total suitable area within each EEZ in order to rank zones by priority. To do so, we need to find the total area of suitable locations within each EEZ.

select suitable cells within West Coast EEZs
```{r}
area_cell <- cellSize(sst_depth, unit = "km")

sst_depth_eez <- area_cell[eez_raster, drop = FALSE]
plot(sst_depth_eez)

eez_raster <- rasterize(west_eez, sst_depth, field="rgn")
plot(eez_raster)


# Crop by new raster
sst_depth_eez <- sst_depth[eez_raster, drop = FALSE]
test_eez <- eez_raster[sst_depth, drop = FALSE]
plot(test_eez)
# Mask raster
# sst_depth <- mask(sst_depth, eez_raster)
plot(sst_depth_eez)

# Zonal: summarize values in raster based on zones in raster
# zonal_oysters <- zonal(sst_depth_eez, eez_raster, fun = "sum", na.rm=TRUE)
# zonal_oysters

```

find area of grid cells
```{r}
area_cell <- cellSize(sst_depth_eez, unit = "km")
plot(area_cell)

eez_raster <- rasterize(west_eez,
                        area_cell,
                        field = 'rgn')

# Initial plot
plot(eez_raster)

# Use zonal algebra to aggregate a grouping variable
eez_suitable <- zonal(x = sst_depth_eez, 
                      z = eez_raster, # Raster representing zones
                      fun = 'sum', # To add up total area
                      na.rm = TRUE)
eez_suitable
# the_answer <- sst_depth * area_cell
# the_answer
# 
# zonal_answer <- zonal(the_answer, eez_raster, fun = "sum", na.rm=TRUE)
# zonal_answer
```

find the total suitable area within each EEZ
hint: it might be helpful to rasterize the EEZ data
```{r}
# Find the area of grid cells using cellSize

suitable_area <- cellSize(x = masked_location, # Selecting suitable locations from above
                 mask = TRUE, # When true, previous NA will carry over
                 unit = 'km') # Selecting km from data

# Initial plot
plot(suitable_area)
# Find the total suitable area within each EEZ

# Rasterize EEZ data
eez_raster <- rasterize(eez,
                        suitable_area, # to this raster
                        field = 'rgn') # Transfer values to each eez region

# Initial plot
plot(eez_raster)

# Use zonal algebra to aggregate a grouping variable
eez_suitable <- zonal(x = suitable_area, 
                      z = eez_raster, # Raster representing zones
                      fun = 'sum', # To add up total area
                      na.rm = TRUE)

# Print suitble area by EEZ
print(paste0("Oyster suitable area in ", eez_suitable$rgn[1], " is ", eez_suitable$area[1], 'km'))
print(paste0("Oyster suitable area in ", eez_suitable$rgn[2], " is ", eez_suitable$area[2], 'km'))
print(paste0("Oyster suitable area in ", eez_suitable$rgn[3], " is ", eez_suitable$area[3], 'km'))
print(paste0("Oyster suitable area in ", eez_suitable$rgn[4], " is ", eez_suitable$area[4], 'km'))
print(paste0("Oyster suitable area in ", eez_suitable$rgn[5], " is ", eez_suitable$area[5], 'km'))



```

